### 2024-12-04 other/datasheet/datasheet_template.qmd
updated datasheet

---
title: "Datasheet for the 'air raids dataset'"
subtitle: "Supplementary document to give context to the data"
author: 
  - Dennis Netchitailo
thanks: "Code and data are available at: https://github.com/dennisnetchitailo/air-raids."
date: "03 December 2024"
date-format: 
abstract: "The UK government required data about the effects of air raids in order to understand the impact on the population and the military. Germany's goal was to weaken the UK through the air enough to allow for an amphibious invasion. It was critical for the UK to understand the effectiveness of Germany's strategy of trying to degrade the UK's morale, economy, and war fighting capabilities."
format: pdf
number-sections: true
bibliography: "C:/Users/Dennis Netchitailo/Documents/air-raids/paper/references.bib"
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to analyze the impact of air raids in the UK during WWII, focusing on casualties and incident patterns. It aimed to provide insights into the frequency and severity of incidents over time.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The UK Ministry of Home Security as part of the Home Security Daily Intelligence Reports
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The ministry was government funded. 
4. *Any other comments?*
    - The dataset made up the vast amount of data collection that was undertaken by the UK government to aid the war effort.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each instance represents an air raid incident, including data on casualties, incidents, time, and location.
2. *How many instances are there in total (of each type, if appropriate)?*
	- There are 32869 instances. 
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset contains all possible instances. It is possible some instances were not recorded, but this would have had to have occurred in areas without human observation (uninhabited areas or the English channel), which would imply a lack of damage to people or infrastructure. 
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Processed features such as year, month, location, casualties, and lethality category.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Yes, the lethality_category variable is used to classify incidents by severity.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Some data points such casualty numbers are unavailable due to incomplete records.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- There is an explicit relationship between years, regions, and casualties.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- There are no data splits used in this analysis. 
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- The errors would relate to incomplete historical records. It wasn't always possible to record a location or a casualty number due to complexity of determining how it is to be measured. 
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is self-contained, but links to the source historical archive for verification. 
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, as the data describes a historical event in the aggregate. No personal details of any kind are in the data. 
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- The dataset does not identify any sub-populations.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- It is indirectly possible. With the coordinates, it is possible to identify current residents at places with incidents, but it would not be possible to identity the people at the incident. However, given another dataset that recorded the location from where a casualty was taken, it would be possible to link an individual to an incident. 
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- The sensitivity of the data is that a current location may be identified as having one been struck during an air raid. This could, for example, influence the sale of a building, if it became known that it had once been struck, due to a fear of its foundation being weakened. 
16. *Any other comments?*
	- The dataset analyzes contains high level data. Even the coordinate data, the most sensitive and identifying of all the data, doesn't cover all individual buildings that were struck.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The dataset was acquired from the datasheet made publicly available by Alex Cookson. 
	- Likely from historical records, government archives, or academic sources. This would involve direct observation of targeted locations, recording 
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Manual digitization from historical documents.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset is exhaustive. 
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The data was collected by police, air raid wardens, and military personnel. This was part of their official duties. 
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data was collected from 1939 to 1945, which matches the timeframe of the data. 
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No. 
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- Third-party government historical archives.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Likely not. 
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Unlikely that consent was obtained. 
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- See above. 
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No. 
12. *Any other comments?*
	-

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Data cleaning, creating derived variables like lethality_category, and handling missing values.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes, available at: https://github.com/tacookson/data/tree/master/britain-bombing-ww2/raw
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- Unknown what software was used. 
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- Be aware of the historical context of the dataset. Avoid misinterpretation. 
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- Avoid use in contexts that trivialize or sensationalize casualty data.
6. *Any other comments?*
	- Be ethical and respectful in using this data for analysis or for interpetation.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- No.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset is freely available on GitHub. 
3. *When will the dataset be distributed?*
	- After review and validation.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Open data license or public domain for historical data.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No. 
7. *Any other comments?*
	-

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The author, Dennis Netchitailo.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- By email: dennis.netchitailo@mail.utoronto.ca
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- The dataset will only be updated to correct errors. 
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- Not applicable for historical data.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- No.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Contact by email.
8. *Any other comments?*
	-

\newpage

# References